Harmony: A Systems Programming Language and Model Checker

### 2022-01-16 paper/paper.tex
Latest

Harmony can be used to model a distributed system in which the
only shared variable is a set of messages representing the network.
From a model checking perspective, it is convenient that message and
timer handling can be performed atomically, as the only interaction
between threads is through the network variable.
A message is sent by atomically adding a message to the network:
<{def send(m): atomically network |= {m}}> (`<{|}>' is the union
operator in Python).
To receive a message, a program can use the Harmony <{when}>
statement to wait for a message.  Typically, however, the program
does not want to receive just any message, but a message that satisfies
certain conditions such as: \emph{wait for the next message that is addressed
to me and has a sequence number larger than 3}.
For this, Harmony supports the <{exists}> clause and allows
<{when}> clauses to be nested:
<{atomically when exists m in network when (m.dst == self) and (m.seq > 3): ...}>
With these constructs, Harmony can be effectively used as a
model checker for distributed algorithms.

>>>>

Harmony can be used to model a distributed system in which the
only shared variable is a set of messages representing the network.
From a model checking perspective, it is convenient that message and
timer handling can be performed atomically, as the only interaction
between threads is through the network variable.
A message is sent by atomically adding a message to the network:
<{def send(m): atomically network |= {m}}> (`<{|}>' is the symbol for
the union operator in Python).
To receive a message, a program can use the Harmony <{when}>
statement to wait for a message.  Typically, however, the program
does not want to receive just any message, but a message that satisfies
certain conditions such as: \emph{wait for the next message that is addressed
to me and has a sequence number larger than 3}.
For this, Harmony supports the <{exists}> clause and allows
<{when}> clauses to be nested:
<{atomically when exists m in network when (m.dst == self) and (m.seq > 3): ...}>
With these constructs, Harmony can be effectively used as a
model checker for distributed algorithms.
